---
title: "Catbond paper by Saeid Bayesian models"
author: "Arpita Chatterjee"
date: '2023-02-25'
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:\\Users\\GSUProf\\Dropbox\\catbond\\Saeid')
knitr::opts_chunk$set(fig.width=6, fig.height=3)
#install.packages("R2OpenBUGS")
##install.packages("CODA")
library(R2OpenBUGS)
library(coda)
library(lattice)

library(stats) # require for quantile function
```

## Model 1 (as in Equation 5.39)

Creating data for Bayesian model


```{r message=FALSE}
x <- c(1290, 2062,    1, 1033,  173,  620, 1832,    1,  551, 2261,  952,   12,  259,  556,
847,  408,  136, 1026,  528,   45, 1108,  219,  635,  341,   32,  548,  303,   35,
223,   49 , 502,   13 , 879,  595 ,  25,  488,  576,  397,  269,  483,  865 , 495,
743,   64,  701,  126,  120 , 988 , 358,   70,  101,  115,    7,   63,  104,  169,
736,   71,  108,   19 , 160,  642,  163,    5,    8,  437,   54,  580,  114,  133,
485,  302 ,  64,  270,  180,  118,  343, 2078,  178,  613,  414,  217,   64 )
data <- list(w=x, m=length(x))
```

Creating initials for the Bayesian model

```{r message=FALSE}
inits <- function(){
  list(xi=c(0.1,0.1))}
```

Let's write the gamma model

```{r message=FALSE}
gammamodel <- function() {
    for (i in 1:m){
     w[i] ~ dgamma(xi[1],xi[2])
    }
   dummy <- 0
   dummy ~ dloglik(phi)
   phi <- log(xi[2])-loggam(xi[1])+((xi[1]-1)*(log(xi[1])-(1/(2*xi[1])))/exp(loggam(xi[1]))-xi[1])
   xi[1] ~ dunif(0.01,1000)
   xi[2] ~ dunif(0.01,1000)

   a <- xi[1]/xi[2] #This line will evaluate E(xi1/xi2|W)
  # b <- (log(xi[1])-(1/(2*xi[1])))-log(xi[2]) #This line will evaluate E(psi(xi1)-log(xi2)|W)
   
   b <- (log(xi[1]+1)-(1/(2*(xi[1]+1)))) - (log(xi[1])-(1/(2*xi[1])))

}
```

Next we'll execute the model

```{r message=FALSE}
bayesout1<- bugs(data, inits, model.file = gammamodel,
               parameters = c("xi", "a", "b"), n.chains = 1, n.burnin=10000, n.iter = 40000,DIC=FALSE, codaPkg=TRUE,digits=3,debug=F)


## Formatting the posterior samples
out <- read.bugs(bayesout1)

```
out[[1]] is a matrix of posterior samples: (1) column 1 gives the samples of $\xi_1/\xi_2$, (2) column 2 gives the samples of $\psi(\xi_1+1)-\psi(\xi_1)$, (3) column 3 gives the sample of $\xi_1$, and (4) column 4 gives the samples of $\xi_2$. 


To obtain the two posterior expected values, $\mathbb{E}^{\mathbb{P}}\left[\frac{\xi_1}{\xi_2} \mid \mathbf{w}\right]$ and $\mathbb{E}^{\mathbb{P}}\left[\psi\left(\xi_1+1\right)-\psi\left(\xi_1\right) \mid \mathbf{w}\right]$, consider the column 1 and column 2 averages respectively. 

We obtain the posterior estimates of $\xi_1$ and $xi_2$ as,

$\hat{\xi_1} =\frac{1}{\mathbb{E}^{\mathbb{P}}\left[\psi\left(\xi_1+1\right)-\psi\left(\xi_1\right) \mid \mathbf{w}\right]}$

$\hat{\xi_2} = \frac{\hat{\xi_1}}{\mathbb{E}^{\mathbb{P}}\left[\frac{\xi_1}{\xi_2} \mid \mathbf{w}\right]}$ and 


```{r message=FALSE}

xi1.hat <- 1/(mean(out[[1]][,2]))
xi2.hat <- xi1.hat/mean(out[[1]][,1])


xi1.hat
xi2.hat

```



To check the model convergence, try running multiple chains. 

```{r message=FALSE}
inits <- function(){
  list(xi=c(0.1,0.1))
  list(xi=c(2,2))
  list(xi=c(0.01,0.01))}

```

```{r message=FALSE}
bayesout_3chains<- bugs(data, inits, model.file = gammamodel,
               parameters = c("xi", "a", "b"), n.chains = 3, n.burnin=5000, n.iter = 15000,DIC=FALSE, codaPkg=TRUE,digits=3,debug=F)


## Formatting the posterior samples
out_3chains <- read.bugs(bayesout_3chains)
```

Convergence diagnostics: 

1. History plot

```{r message=FALSE}
#library(coda)
#library(lattice)

xyplot(out_3chains)
```
2. Density plot:

```{r message=FALSE}
densityplot(out_3chains)
```
3. ACF plot

```{r message=FALSE}
acfplot(out_3chains)
```




## Model 2 (Mixture of Exponential)


```{r message=FALSE}
mixexp <- function() {
for (i in 1:m){
gr[i] ~dcat(p[])
w[i] ~ dexp(lambda[gr[i]])
}

p[1:2]~ddirich(alpha[])
for(k in 1:2){
lambda[k] ~ dgamma(tau[k],psi[k])
}

}
```

Creating data for Bayesian model


```{r message=FALSE}
data <- list(w=x, m=length(x),tau=c(0.01,0.01), psi=c(0.01,0.01), alpha=c(0.5,0.5))
```

Creating initials for the Bayesian model

```{r message=FALSE}
inits <- function(){
  list(lambda=c(0.1,0.1))}
```

Let's execute the model

```{r message=FALSE}
bayesout_mix<- bugs(data, inits, model.file = mixexp,
               parameters = c("lambda"), n.chains = 1, n.burnin=10000, n.iter = 40000,DIC=FALSE, codaPkg=TRUE,digits=3,debug=F)

out_mixexp <- read.bugs(bayesout_mix)
```

Let's obtain the posterior estimates under the squared error loss


```{r message=FALSE}
bayes.est <- colMeans(out_mixexp[[1]]) 
bayes.est


#95% credible interval 
cred.int <- cbind(quantile(out_mixexp[[1]][,1], prob=c(0.025,0.975)), quantile(out_mixexp[[1]][,2], prob=c(0.025,0.975)))
```
Convergence diagnostics: 

1. History plot

```{r message=FALSE}
#library(coda)
#library(lattice)

xyplot(out_mixexp)
```

The model doesn't converge well. 



